---
title: "Welcome to Shiny"
author: "Erik Ländström"
date: "14 December 2018"
output: 
  html_document:
    keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Examples

Shiny has 11 built in examples that demonstrate how Shiny works.

The __Hello Shiny__ example plots a histrogram of R's `faithful`
dataset.

```{r hello_shiny}
library(shiny)
runExample("01_hello")
```

# Structure of a Shiny App

Shiny apps are contained in a single script called `app.R`. The script `app.R` lives in a directory and the app can be run with `runApp("newdir").

`app.R` has three components:

* a user interface object

* a server function

* a call to the `shinyApp` function

The user interface (`ui`) object controls the layout and appearance of your app. The `server` function contains the instructions that your computer needs to build your app. Finally the `shinyApp` function creates Shiny app objects from an explicit UI/server pair.

## __ui__

```{r ui}
# Define UI for app that draws a histogram ----
ui <- fluidPage(

  # App title ----
  titlePanel("Hello Shiny!"),

  # Sidebar layout with input and output definitions ----
  sidebarLayout(

    # Sidebar panel for inputs ----
    sidebarPanel(

      # Input: Slider for the number of bins ----
      sliderInput(inputId = "bins",
                  label = "Number of bins:",
                  min = 1,
                  max = 50,
                  value = 30)

    ),

    # Main panel for displaying outputs ----
    mainPanel(

      # Output: Histogram ----
      plotOutput(outputId = "distPlot")

    )
  )
)
```

## __server__

```{r server}
# Define server logic required to draw a histogram ----
server <- function(input, output) {

  # Histogram of the Old Faithful Geyser Data ----
  # with requested number of bins
  # This expression that generates a histogram is wrapped in a call
  # to renderPlot to indicate that:
  #
  # 1. It is "reactive" and therefore should be automatically
  #    re-executed when inputs (input$bins) change
  # 2. Its output type is a plot
  output$distPlot <- renderPlot({

    x    <- faithful$waiting
    bins <- seq(min(x), max(x), length.out = input$bins + 1)

    hist(x, breaks = bins, col = "#75AADB", border = "white",
         xlab = "Waiting time to next eruption (in mins)",
         main = "Histogram of waiting times")

    })

}
```

At one level, the __Hello Shiny__ `server` function is very simple. The script does some calculations and then plots a histogram with the requested number of bins.

However, you’ll also notice that most of the script is wrapped in a call to `renderPlot`. The comment above the function explains a bit about this, but if you find it confusing, don’t worry. We’ll cover this concept in much more detail soon.

Play with the __Hello Shiny__ app and review the source code. Try to develop a feel for how the app works. But before you do so, note that in your `app.R` file you will need to start with loading the Shiny package and end with a call to shinyApp:

```{r example}
library(shiny)

# See above for the definitions of ui and server
# ui <- ...

 #server <- ...

shinyApp(ui = ui, server = server)
```

Your R session will be busy while the __Hello__ Shiny app is active, so you will not be able to run any R commands. R is monitoring the app and executing the app’s reactions. To get your R session back, hit escape or click the stop sign icon (found in the upper right corner of the RStudio console panel).

# Running an App

Every Shiny app has the same structure: an `app.R` file that contains `ui` and `server`. You can create a Shiny app by making a new directory and saving an `app.R` file inside it. It is recommended that each app will live in its own unique directory.

You can run a Shiny app by giving the name of its directory to the function `runApp`. For example if your Shiny app is in a directory called `my_app`, run it with the following code:

```{r running, eval = FALSE}
library(shiny)

runApp("my_app")
```

Note: `runApp` is similar to read.csv, read.table, and many other functions in R. The first argument of `runApp` is the filepath from your working directory to the app’s directory. The code above assumes that the app directory is in your working directory. In this case, the filepath is just the name of the directory.

# My turn

Create a new directory named `App-1` in your working directory. Copy and paste the app.R script above into your directory. 

```{r run_app}
runApp("App-1")
```

Run app in "showcase mode" to show the script alongside the app.

```{r showcase}
runApp("App-1", display.mode = "showcase")
```

# More examples

```{r, eval = FALSE}
runExample("01_hello")      # a histogram
runExample("02_text")       # tables and data frames
runExample("03_reactivity") # a reactive expression
runExample("04_mpg")        # global variables
runExample("05_sliders")    # slider bars
runExample("06_tabsets")    # tabbed panels
runExample("07_widgets")    # help text and submit buttons
runExample("08_html")       # Shiny app built from HTML
runExample("09_upload")     # file upload wizard
runExample("10_download")   # file download wizard
runExample("11_timer")      # an automated timer
```

Lesson taken from [shiny.rstudio.com](https://shiny.rstudio.com/tutorial/written-tutorial/lesson1/)































